---
title: "Exercise of Data Transform"
author: "Jamia Begum"
date: "2022-10-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Exercises (data transform) :
1. Using flights dataset, find all flights that
a) Had an arrival delay of two or more hours
b)Flew to Houston (IAH or HOU) 
c) Arrived more than two hours late, but didn’t leave late
d) Were delayed by at least an hour, but made up over 30 minutes in flight


```{r}
library(tidyverse)
library("nycflights13")
library("dplyr")
flights
filter(flights, arr_delay >= 120)
select(flights,dest)
filter(flights, dest == "IAH" | dest == "HOU") 
filter(flights, arr_delay > 120, dep_delay <= 0)
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
```


2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?


between(x, left, right) gives values in a numeric vector bounded in a specified range (left, right) 

Using it in the previous questions does not simplify the codes here. 
```{r}
filter(flights, between(arr_delay, 120, Inf) )
filter(flights, between(arr_delay, 120, Inf)& between(dep_delay,-Inf,0) )
filter(flights, between(dep_delay, 60, Inf)& between(dep_delay- arr_delay,31,Inf) )
```

3. Sort flights to find the fastest flights.

fastest flights mean having highest average speed calculated by distance/air time
```{r}
arrange(flights, desc(distance / air_time))

```

4. Create a new data frame having variables with the dep string.


```{r}
dep_str=select(flights, contains("dep"))
dep_str
```

5. Create a new data frame having the hour and minute of
depature (Hint: information is in the variable dep_time with format HHMM or HMM. Use %/% or %% when appropriate)

we need to split hour digits and minute digits from dep_time.
## %/% gives quotient so gives the hour digits here
##  %%  gives remainder so gives the minute digits here

```{r}
dp_flights<-mutate(flights,dep_hour=dep_time %/% 100, dep_min = dep_time %% 100)
select(dp_flights,dep_time,dep_hour,dep_min)
```



6. Create a summary of each airline (variable carrier) describing the total number of flights, the average, median, IQR of arr_delay
```{r}
flights %>%
  group_by(carrier) %>% summarise( count = n(), mean_delay = mean(arr_delay, na.rm = TRUE)
  ,median_delay = median(arr_delay, na.rm = TRUE),
  IQR_delay = IQR(arr_delay, na.rm = TRUE))
```


