---
title: "Exercise for ggplot2"
author: "Jamia Begum"
date: "2022-10-14"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Exercise:1
Use what you’ve learned to improve the visualisation of the departure times of cancelled vs. non-cancelled flights (NOTE: missing values in the dep_time variable indicate that the flight was cancelled).

```{r}
library(tidyverse)
library(dplyr)
library(nycflights13) #to get flights tibble
flights
f<-mutate(flights,cancel=is.na(dep_time))
#true=canceled flights
#false=Non-canceled flights
pl<-select(f,dep_time,cancel)
pl
library(ggplot2)

       ggplot(pl,aes(x=cancel,fill=cancel)) +ggtitle("Canceled flights vs Non-canceled flights") +
         xlab("catergorized flights") +
         geom_bar()+ geom_text(stat='count',aes(label=after_stat(count)), vjust=-0.5)
```

#Exercise:2
a)What variable in the diamonds dataset is most important for predicting the price of a diamond? 
Here,R package corrplot should be used to visualize the correlation matrix that shows the linear correlations between variables.

```{r}

library(corrplot)
diamonds
M <- cor(diamonds[,c(1,5,6,7,8,9,10)])
M
corrplot(M, method = 'number')
ggplot(data = diamonds, mapping = aes(x = carat, y = price)) +
  geom_boxplot(mapping = aes(group = cut_width(carat,0.2)))+
  ggtitle("Carat vs Price")


```
From the correlation Matrix, we see that "carat" has the highest correlation coefficient related to "price" .It means "carat" has a fairly strong positive relationship with "price".Also, the boxplots between "carat" and "price" shows that in most of the cases the price is increasing as the "carat" sizes go up.So, "carat" is most important for predicting the price of a diamond.


b)How is that variable("carat") correlated with cut?
How is that variable correlated with cut? Why does the combination of those two relationships lead to lower quality diamonds being more expensive?. NOTE: variable cut describes the quality (see ?diamonds).

Here, "carat" is a continuous variable but "cut" is a categorical variable. To visualize the correlation them ,it's better to use boxplots

```{r}
ggplot(diamonds, aes(x = cut, y = carat)) +
  geom_boxplot()+ggtitle("Carat vs Cut")
```
we see that the relation fluctuates so much between "cut" and "carat". Here, some low quality cuts (Fair) are having larger "carat" ,thus higher prices whereas good and ideal  quality diamonds are having less "carat" and low prices.



#Exercise-4
Instead of summarising the conditional distribution with a boxplot, you could use a frequency polygon.

Here, "freqpoly" plot gives the conditional distribution by showing which cut has the highest density with the price.

```{r}
ggplot(data = diamonds, mapping = aes(x = price)) +
  geom_freqpoly(mapping = aes(color = cut), binwidth = 5000)+
  ggtitle("Freqpoly Plot of Carat vs Cut")
```


What do you need to consider when using cut_width() vs cut_number()? 

we can use the functions "cut_width()" and "cut_number()" in "geom_freqpoly" to split a variable into groups.
#For "cut_width()",we need to specify the widths and then number of groups will be calculated automatically.
#For "cut_number()",we need to specify the number of groups, the the width of each group will be adjusted automatically.

How does that impact a visualisation of the 2d distribution of carat and price (diamonds dataset)?

#using "cut_width()" to split "carat" variable into groups and then doing the freqpoly plot to see the density of the price for each group of carat

```{r}
ggplot(data = diamonds,
  mapping = aes(color = cut_width(carat, .5, boundary = 0), x = price)) +
  geom_freqpoly() +
  labs(x = "Price", color = "Carat")+ggtitle("Carat vs price(cut_width)")
```

#using "cut_number()" to split "carat" variable into groups and then doing the freqpoly plot to see the density of the price for each group of carat

```{r}
ggplot(data = diamonds,
  mapping = aes( x = price)) +
  geom_freqpoly( mapping = aes(color = cut_number(carat, 6))) +
  labs(x = "Price", color = "Carat")+ggtitle("Carat vs Price(cut_number)")
```



#Exercise-6
Download genome data:(https://raw.githubusercontent.com/isglobalbrge/Master_Modelling/main/data/genome.txt) into your computer and load it into RStudio by using read_delim function (NOTE: data are tab-delimited).

```{r}
setwd("F:/INTERMATH/intermath 2021-2023/spain/DV/R for data Science")
genome<-read_delim('genome.txt')
genome

#6.1)Which is the expected value of Log.R.Ratio and B.Allel.Freq for each chromosome? (show the R code to get your answer)

genome %>%
  group_by(Chr) %>% 
summarise(mean_log = mean(Log.R.Ratio, na.rm = TRUE)
          ,mean_B.Allele = mean( B.Allele.Freq, na.rm = TRUE))



#6.2: Create a facet plot that represent the Log.R.Ratio for each chromosome

ggplot(data = genome, aes(Position,Log.R.Ratio)) +
  geom_line(color = "blue", size = 1) +
  labs(title = "Facet plot of Log.R.Ratio",
       y = "Log.R.Ratio", x = "Position") + 
  facet_wrap(~ Chr)

#6.3: Create a facet plot of B.Allele.Freq for chromosomes 1, 2, 3, . . . , 6 drawing B.Allele.Freq information in red color.

ggplot(data = genome, aes(Position,B.Allele.Freq)) +
  geom_point(color="red",size = .5) + 
  labs(title = "Facet plot of B.Allele.Freq",
       x = "position", y = "B.Allele.Freq") + 
  facet_wrap(~ Chr)
```











Exercise: 3 Visualize the number of flights of each airline by month.

```{r}
library(nycflights13)
print(count(flights,carrier,month) ,n=180)

ggplot(data = flights, mapping = aes(x = factor(month),fill=factor(month))) +
  geom_bar() +
  facet_wrap(~ carrier, scales= "free")
```

